---
title: "NOAA-AFSC-NBS-sample-map"
author: "Kimberly Ledger"
date: "2024-02-14"
output: html_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

load libraries
```{r}
library(tidyverse)
#library(RColorBrewer)
#library(ggOceanMaps)
#library(ggspatial)
```

get data - i'm pretty sure eDNA samples were taken along with the CTD casts, so i'll extract the location information from those
```{r}
#eDNA samples
samples <- read.csv("~/NBS_eDNA_sample_map/data/NBS_eDNA_extracts_as_of_20240214.csv") %>%
  filter(sample_type == "sample") #for this i just care about samples, not any of the extraction blanks or field blanks

#NBS2021 event data
nw2101 <- read.csv("~/NBS_eDNA_sample_map/data/NBS_nbs21_event.csv") %>%
  filter(GEAR == "CTD") %>%
  select(SURVEY, STATION, EQ_TIME, EQ_LATITUDE, EQ_LONGITUDE) %>%
  mutate(SURVEY = as.factor(SURVEY))%>%
  mutate(STATION = as.factor(STATION))

#NBS2022 event data
nw2202 <- read.csv("~/NBS_eDNA_sample_map/data/NBS_nw2202_event.csv") %>%
  filter(GEAR == "CTD") %>%
  select(SURVEY, STATION, EQ_TIME, EQ_LATITUDE, EQ_LONGITUDE) %>%
  mutate(SURVEY = as.factor(SURVEY))%>%
  mutate(STATION = as.factor(STATION))

#NBS2023 event data
nw2301 <- read.csv("~/NBS_eDNA_sample_map/data/NBS_nw2301_event.csv") %>%
  filter(GEAR == "CTD") %>%
  select(SURVEY, STATION, EQ_TIME, EQ_LATITUDE, EQ_LONGITUDE) %>%
  mutate(SURVEY = as.factor(SURVEY))%>%
  mutate(STATION = as.factor(STATION))

#looks like the stationID is missing from CTD 5 so i'll fix that manually 
nw2301[44,2] <- 5
```

to join location data to eDNA samples, add a column to samples with the survey ID 
```{r}
samples <- samples %>%
  mutate(SURVEY = ifelse(collection_year == "2021", 202101, "NA"),
         SURVEY = ifelse(collection_year == "2022", 202202, SURVEY),
         SURVEY = ifelse(collection_year == "2023", 202301, SURVEY)) %>%
  rename(STATION = location1) %>%
  select(extraction_ID, collection_year, STATION, depth, SURVEY)  ## clean up the df to make it easier to look at.. 

samples$SURVEY <- as.factor(samples$SURVEY)
samples$STATION <- as.factor(samples$STATION)
samples$collection_year <- as.factor(samples$collection_year)
samples$depth <- as.character(samples$depth)
```

missing lat/long from st43 in 2023
```{r}
st43_nw2301 <- read.csv("~/NBS_eDNA_sample_map/data/NBS_nw2301_event.csv") %>%
  filter(GEAR == "Bongo") %>%
  filter(STATION == 43) %>%
  select(SURVEY, STATION, EQ_TIME, EQ_LATITUDE, EQ_LONGITUDE) %>%
  mutate(SURVEY = as.factor(SURVEY)) %>%
  mutate(STATION = as.factor(STATION))
```


also add in the SBS samples 
```{r}
#SBS2022 - DY22 event data 
DY22 <- read.csv("~/NBS_eDNA_sample_map/data/DY22-08Event.csv") %>%
  filter(SampleYear == "2022") %>%
  filter(CruiseID == "8DY22") %>%
  dplyr::select(CruiseID, StationNumber, EQ.Latitude, EQ.Longitude) %>%
  rename(SURVEY = CruiseID) %>%
  rename(STATION = StationNumber) %>%
  rename(EQ_LATITUDE = EQ.Latitude) %>%
  rename(EQ_LONGITUDE = EQ.Longitude) %>%
  mutate(STATION = as.factor(STATION)) %>%
  mutate(SURVEY = as.factor(SURVEY))

#SBS eDNA samples 
SBS_samples <- read.csv("~/NBS_eDNA_sample_map/data/SBS_eDNA_extracts_as_of_20240215.csv") %>%
  filter(sample_type == "sample") %>% #for this i just care about samples, not any of the extraction blanks or field blanks
  mutate(SURVEY = ifelse(location3 == "PAN21", "PAN21", "8DY22")) %>%
  mutate(collection_year = ifelse(SURVEY == "8DY22", 2022, collection_year)) %>%
  mutate(collection_year = as.factor(collection_year)) %>%
  mutate(depth = as.character(depth)) %>%
  select(extraction_ID, collection_year, location1, depth, SURVEY, longitude, latitude) %>% ## clean up the df to make it easier to look at.. 
  rename(STATION = location1) %>%
  mutate(STATION = as.factor(STATION))
```


filter extraction to just keep unique sampling events (i.e. get rid of the three field reps for most samples) 
and join the lat/long 
```{r}
surveys <- nw2101 %>%
  bind_rows(nw2202) %>%
  bind_rows(nw2301) %>%
  bind_rows(st43_nw2301) %>%
  bind_rows(DY22)

unique_samples <- samples %>%
  bind_rows(SBS_samples) %>%
  select(!extraction_ID) %>%
  unique() %>%
  left_join(surveys, by = c("SURVEY", "STATION")) %>%
  mutate(EQ_LATITUDE = ifelse(!is.na(latitude), latitude, EQ_LATITUDE)) %>%
  mutate(EQ_LONGITUDE = ifelse(!is.na(longitude), longitude, EQ_LONGITUDE)) %>%
  select(!latitude) %>%
  select(!longitude)

#unique_samples$collection_year <- as.factor(unique_samples$collection_year)
unique_samples$EQ_LATITUDE <- as.numeric(unique_samples$EQ_LATITUDE)
unique_samples$EQ_LONGITUDE <- as.numeric(unique_samples$EQ_LONGITUDE)
#unique_samples$depth <- as.factor(unique_samples$depth)
```

make visual of sampling locations
```{r}
map_with_sites <- basemap(c(-177, -159, 58, 67), bathymetry = FALSE, rotate = TRUE) + 
  ggspatial::geom_spatial_point(data = unique_samples, aes(x = EQ_LONGITUDE, y = EQ_LATITUDE, color = collection_year), size = 1) + 
  labs(color = "Year")
map_with_sites
```

try a different plotting package 
```{r}
library(rnaturalearth)
library(rnaturalearthdata)
world <- ne_countries(scale = "medium", returnclass = "sf")
```


# Cruise Map

```{r,echo=FALSE,warning=FALSE, fig.height= 8, fig.width= 12}
min_lat <- min(unique_samples$EQ_LATITUDE, na.rm = T)
max_lat <- max(unique_samples$EQ_LATITUDE, na.rm = T)

min_lon <- min(unique_samples$EQ_LONGITUDE, na.rm = T)
max_lon <- max(unique_samples$EQ_LONGITUDE, na.rm = T)

unique_samples$collection_year <- as.factor(unique_samples$collection_year)

#bin the 30-60m depth samples to make a figure with fewer grids
unique_samples <- unique_samples %>%
  #mutate(depth2 = as.character(depth)) %>%
  mutate(depth2 = ifelse(depth == 30, "30 to 60", depth),
         depth2 = ifelse(depth == 40, "30 to 60", depth2),
         depth2 = ifelse(depth == 50, "30 to 60", depth2),
         depth2 = ifelse(depth == 60, "30 to 60", depth2))


BeringSea_plot <- ggplot(data = world) +
    geom_sf() +
    geom_point(data = unique_samples, aes(x = EQ_LONGITUDE, y = EQ_LATITUDE, color= SURVEY)) + 
  coord_sf(xlim = c(min_lon-2, max_lon+2), ylim = c(min_lat-1, max_lat+1), expand = FALSE) +
  theme_bw() +xlab("Longitude") +ylab("Latitude") +
  facet_grid(depth2~collection_year) + 
  theme(axis.text.x = element_text(angle = 90))

BeringSea_plot

ggsave(plot= BeringSea_plot, 
      filename = ("~/NBS_eDNA_sample_map/figures/BeringSea_plot.png"),
      width=12,
      height = 8,
      dpi = 300,
      units = c("in"))
```
